%!TEX root = ../dissertation.tex

\chapter{User Study Protocol}
\label{appendix:protocol}

\section{Motivation}
%
By conducting this first study, we intend to:
%
\begin{itemize}
  \item Conclude if there is any chance that cultural behaviours influence the user's color perception.
  \item Realize which color mixtures are more easily perceived by humans.
  \item Understand if, by using color, it is possible to clearly and easily convey information.
  This can be particularly interesting and useful when visualizing graphs or maps.
  \item Conclude if a person is capable of, not only building a mental color mixture model, but
  also deconstructing mixtures into their basic components.
\end{itemize}
%
\section{User Profiling Phase}
%
This study is anonymous and should take you up to 15 minutes. Please, answer the following answers accordingly.
%
\section{Testing Calibration Phase}
%
In this step, it's going to be presented to you a set of images. You should tune you screen definitions, in
order to answer the questions, keeping them until the end of this study. \par
%
Please, follow the steps below indicated and answer the questions.
%
\begin{enumerate}
  \item If possible, adjust your room lights for a comfortable usage of your device.
  \item Avoid reflections on your screen, by diverting the screen from direct sources of light. This step is important,
  since light reflections can affect visualization of images.
  \item To adjust the \textbf{Black Point} of your screen, define the \ul{Contrast} and \ul{Brightness} of your screen to their maximum.
  \item After Step 3, gradually reduce \textbf{Brightness} value of your screen, in order to correctly distinguish the squares of each image below [calibration squares images].
  \item If possible, define the \textbf{Color Temperature} of your screen to 6500 Kelvin Degrees.
  \item You are now ready to answer the following questions!
\end{enumerate} \par
%
\ul{NOTE:} These 6 steps are only available to the Online Users, since the Laboratory Users
do not need to perform these steps as the LCD display is already calibrated.
%
\section{Testing Color Vision Deficiences Phase}
%
This is the Color Vision Deficiencies Test. \\
%
In this step, it is going to be presented six plates with a colored pattern. Your job is to
identify the number present in each plate, typing it down in the text box below. According
to your answer, this test will inform us if you have any type of color vision deficiency which
may undermine the job of color detection. \\
%
\section{Core Test Phase}
%
Choose the Resulting color which you believe it is the result of mixing the First and Second color, by adjusting the slider below the Resulting color. \\
%
Choose two colors with which you can achieve the Resulting color, by adjusting the sliders below each color. \par
%
\ul{NOTE:} These instructions appear alternately, depending on the type of question which is shown.
%
